\documentclass[13pt, letterpaper, oneside]{book}
\usepackage{graphicx}

\usepackage[driver=xetex,paperwidth=8.5in,paperheight=11in,left=1.2in,
right=1in,top=1.3in, bottom=1.4in]{geometry}
\usepackage[no-math]{fontspec}
\usepackage{newfloat}
\usepackage{sectsty, tikz, color, pgfplots}
\usetikzlibrary{shapes,arrows}
\usepackage{multicol}
\usepackage{amsmath, amssymb, amsfonts,  titlesec}
\usepackage[urw-garamond]{mathdesign}
\usepackage{fancyhdr, booktabs, longtable}
%\usepackage[font=small,format=plain,labelfont=it,textfont=it]{caption}
\usepackage{caption,subcaption}
\usepackage{listings}
\usepackage{algpseudocode, algorithm,setspace}
% \usepackage[T1]{fontenc} 
\usepackage{enumitem,verbatim,natbib}

 \DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ } 
\include{typography}

\usepackage[pdfpagelayout=TwoPageRight, hidelinks]{hyperref}

\begin{document}
%\begin{comment}
\frontmatter

\include{coverpage}
%\include{firstpage}

\tableofcontents
%\end{comment}
\pagestyle{fancy}
\mainmatter
\pagebreak
\vskip 4em

- Introduction (stays the same)
- Background MIP/CP, Literature Review, Description of Malapert's approach
  - What are we trying to improve upon?
- Progress to Date; outline what was done and why.
- Experimental results
- Future work
- Citations
- TOC, List of Figures and Tables, Title Page required

\chapter{Introduction}
This paper discusses three different approaches, and several variations on them,
to solving the problem of scheduling non-identical jobs on a batch processing
machine. Batch processing machines, for the purposes of this paper, can process
multiple, non-identical jobs simultaneously -- but all jobs must be loaded into
and unloaded from the machine at once, which introduces a considerable twist on
the ``simple parallel resources'' known from typical example problems in
existing literature.

The machines in question represents real-life resources like autoclaves or
ovens, which can process multiple items at a time, but often cannot be opened at
random -- in fact, such machines often need to wait for the largest item in the
batch to be done before the next batch can be inserted.

Malapert proposed a custom constraint programming algorithm consisting of a set of
filtering rules to solve the problem. He achieved considerably better speeds
than with a simple mixed-integer model. In this paper, we present 1) an
improvement to his MIP model, 2) a CP model and 3) a decomposition approach to
``divide and conquer'' the problem.

\section{Problem definition}
We describe the problem as follows: assume we are given a set of jobs $J$, each of
which has a processing time (or ``length'') $p_j$ and a size (or ``capacity
requirement'') $s_j$ -- this notation follows Malapert's paper. Each job also
has a due date $d_j$. The machine is characterized by its capacity $b$, and in
every batch, the jobs' summed sizes must not exceed this number. All values are
integer.

The machine can only process one batch $k$ of jobs at a time, and batches always
take as long as the longest job in the batch (i.e. $P_k = \max_{j \in k}(p_j)$).
Our objective is to minimize the lateness $L$ of the latest job in $J$, where
$L$ is the difference between the job's completion time $C_j$ and its due date
$d_j$ -- in formal terms, \textit{min.} $\Lmax = \max_j(C_j - d_j)$. The job's
completion time, however, is the completion time of the batch, which in turn
finishes with its \textit{longest} job as stated above.

Malapert uses the format established by Graham et al. \cite{graham_notation} to
summarize the problem as $1|\textit{p-batch}; b < n;
\textit{non-identical}|\Lmax$, where $\textit{p-batch};b<n$ represents the
parallel-batch nature and the finite capacity of the resource. A simpler version
with identical job sizes was shown to be strongly NP-hard by Brucker et al.
\cite{brucker}, and this problem, then, is no less difficult.

It helps to visualize the jobs before delving into the technicalities of
scheduling them. Figure \ref{fig:intro_tetris} shows a solution to a sample
problem with eight jobs and a resource with capacity $b = 20$.

\input{figures/intro_tetris.tex}
\section{Organization of this paper}
After reviewing some of the most relevant publications on both general MIP/CP models and
batch scheduling problems, we first describe Malapert's original MIP model in
section \ref{sec:malapertmipmodel}. We then present possible improvements to the
model in \ref{sec:improvedmipmodel}. Section \ref{sec:cpmodel} introduces a CP formulation of the same
problem. Sections \ref{sec:mipdecomp} and \ref{sec:cpdecomp} describe a
decomposition approach.

An empirical comparison of the methods and a discussion of the results follow in
sections \ref{sec:results} and \ref{sec:discussion}. Ideas for future work are
listed in \ref{sec:futurework}.



\chapter{Literature Review}
% citet{Azizoglu} results in [2002]
% citep{Azizoglu} results in [Azizoglu et al., 2002]
% citep*{Azizoglu} results in [Azizoglu and Miller, 2002]

Papers I have read so far:
\section{General MIP and CP}
  \begin{alist}
    \item{Fundamentals of mathematical programming}
    \item{Benders}
  \end{alist}
\section{Batch processing}
\begin{alist}
    \item{\citep{Azizoglu}}
    \item{Dupont}
    \item{Sabouni}
    \end{alist}
\subsection{MIP}
\begin{alist}
    \item{Grossmann}
    \end{alist}
\subsection{CP}
\begin{alist}
    \item{Baptiste}
    \item{Malapert (how does he fit into the picture?)}
\end{alist}
\chapter{Modelling the problem}

\input{mipmodel.tex}

\input{cpmodel.tex}
\input{decomp.tex}

\input{testresults.tex}
\input{discussion.tex}
\input{futurework.tex}


\begin{comment}
\chapter{My solution}
\section{MIP formulation improvements}
\section{CP formulation improvements}

\chapter{Discussion}


\end{comment}
\pagebreak

\bibliographystyle{plainnat}
\bibliography{bibliography}{}
\vskip 4em

\end{document}

