\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage[pdftex]{graphicx,xcolor} %\pdfcompresslevel=9  %\pdfimageresolution=600
\newcommand{\newrgbcolor}[2]{\definecolor{#1}{rgb}{#2}}
%%%% listings, env  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{listings}
\definecolor{Brown}{cmyk}{0,0.81,1,0.60}
\definecolor{OliveGreen}{cmyk}{0.64,0,0.95,0.40}
\definecolor{CadetBlue}{cmyk}{0.62,0.57,0.23,0}
\lstset{language=java,frame=ltrb,framesep=5pt,
    numbers=none,
    tabsize=3,
    breaklines=true,
    aboveskip=1ex,
    belowskip=1ex,
    basicstyle=\small\ttfamily,
    backgroundcolor=\color{gray!25},
    columns=fullflexible,
    keywordstyle=\ttfamily\color{OliveGreen},
    identifierstyle=\ttfamily\color{CadetBlue}\bfseries, 
    commentstyle=\color{Brown},
    stringstyle=\ttfamily,
    showstringspaces=true}

\newcommand{\mylst}[1]{\lstinline|#1|}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{Releasing CHOCO 2}
\author{Charles Prud'homme}
%\date{}                                           % Activate to display a given date or no date

\begin{document}
\maketitle

This document requires the correct installation of :
\begin{itemize}
\item Java SE Runtime Environment 6 (or greater),
\item \textcolor{red}{\Huge $<<<<$ Maven 2 $>>>>$},
\item a Subversion client.
\end{itemize}

Feel free to report any errors occurring during the execution of this steps.

\section{Check-out source code}

To check out the source code, run the following comand in a terminal:
\begin{lstlisting}
$>svn co https://choco.svn.sourceforge.net/svnroot/choco/trunk trunk
\end{lstlisting}

\vspace{.5cm}
Go in 'trunk' directory and install the project, using maven2 and the pom.xml:
\begin{lstlisting}
$>mvn clean install -DskipTests
\end{lstlisting}
the option \mylst{-DskipTests} is recommended, as tests are run just after. 

\vspace{.5cm}
Run the tests over the entire project (and take a coffee):
\begin{lstlisting}
$>mvn test -DtestFailureIgnore=true
\end{lstlisting}
The \mylst{-DtestFailureIgnore=true} option avoids hard stop on a failed test. 

At that point, you should consider the failed tests (if any) and fix them if necessary. 

\vspace{.5cm}
Make sure the \mylst{./src/changes/changes.xml} file has been correctly updated since the last release. The changes html file is created during the \mylst{mvn:site} command. 

\vspace{.5cm}
Prepare and perform release on SCM repository (don't forget to mount the release directory):
\begin{lstlisting}
$>mvn release:prepare -DpreparationGoals=clean -Dusername=sf_log -Dpassword=sf_pwd
$>mvn release:perform -DpreparationGoals=clean
\end{lstlisting}
At that point the source repository is up-to-date. Now the released files (ZIP and TAR.GZ) can be created on the newly tagged version.

\vspace{.5cm}
Run the following commands (and take another coffee) to produce the final ZIP and TAR.GZ files:
\begin{lstlisting}
$>mvn clean
$>mvn javadoc:aggregate
$>mvn package -DskipTests
$>mvn assembly:assembly -DskipTests
\end{lstlisting}

\vspace{.5cm}
Checkout the last release version (ex: 2.1.2), then update sourceforge website via SFTP:
\begin{lstlisting}
username,choco@frs.sourceforge.net
/home/frs/project/c/ch/choco/choco/
\end{lstlisting}

%\subsection{}



\end{document}  