%!TEX root = ../content-tut.tex
%\part{getting started}
\label{gettingstarted}
\hypertarget{gettingstarted}{}

\chapter{Getting started: welcome to Choco}\label{gettingstarted:gettingstarted:welcometochoco}\hypertarget{gettingstarted:gettingstarted:welcometochoco}{}
%This introduction covers the basics of writing a program in Choco

%Choco is a java library for constraint satisfaction problems (CSP), constraint programming (CP) and explanation-based constraint solving (e-CP). It is built on a event-based propagation mechanism with backtrackable structures. 

\section{Before starting}\label{gettingstarted:beforestarting}\hypertarget{gettingstarted:beforestarting}{}

Before doing anything, you have to be sure that 
\begin{itemize}
	\item you have at least \href{http://java.sun.com/javase/6/}{Java6} installed on your environment.
	\item you have a IDE (like \href{http://www.jetbrains.com/idea/}{IntelliJ IDEA} or \href{http://www.eclipse.org/}{Eclipse}).
\end{itemize}

To install Java6 or your IDE, please refer to its specific documentation. We now assume that you have the previously defined environment.

You need to create a \textbf{New Project...} on your favorite IDE (\href{http://www.jetbrains.com/idea/training/demos.html}{create a new project on IntelliJ}, \href{https://eclipse-tutorial.dev.java.net/eclipse-tutorial/part1.html}{create a new project on Eclipse}). Our project name is \emph{ChocoProgram}.
Create a new class, named \emph{MyFirstChocoProgram}, with a main method.
\begin{lstlisting}
	public class MyFirstChocoProgram {
	
	    public static void main(String[] args) {
	        
	    }
	}
\end{lstlisting}

\section{Download Choco}\label{gettingstarted:downloadchoco}\hypertarget{gettingstarted:downloadchoco}{}
Now, before doing anything else, you need to download the last stable version of Choco. See the \href{http://www.emn.fr/z-info/choco-solver/choco-download.html}{download page}. 
Once you have download choco, you need to add it to the classpath of your project.

Now you are ready to create you first Choco program.

If you want a short introduction on what is constraint programming, you can find some informations in the \href{http://www.emn.fr/z-info/choco-solver/choco-documentation.html}{Documentation of choco}.\\
When you feel ready, solve your own problem! And if you need more tries, please take a look at the \hyperlink{exercises}{exercises}. %and \hyperlink{examples}{examples}. 

\chapter{First Example: Magic square}\label{gettingstarted:firstexample:magicsquare}\hypertarget{gettingstarted:firstexample:magicsquare}{}
A simple magic square of order 3 can be seen as the ``Hello world!'' program in Choco. First of all, we need to agree on the definition of a magic square of order 3.
\href{http://en.wikipedia.org/wiki/Magic_square}{Wikipedia} tells us that :
\begin{myquote}
A \textbf{magic square} of order $n$ is an arrangement of $n^2$ numbers, usually distinct integers, in a square, such that the $n$ numbers in all rows, all columns, and both diagonals sum to the same constant. A normal magic square contains the integers from 1 to $n^2$.
\end{myquote}

So we are going to solve a problem where unknows are cells value, knowing that each cell can take its value between 1 and $n^2$, is different from the others and columns, diagonals and rows are equal to the same constant M (which is equal to $n * (n^2 + 1) / 2$).

We have the definition, let see how to add some Choco in it.

\section{First, the model}\label{gettingstarted:first,themodel}\hypertarget{gettingstarted:first,themodel}{}
To define our problem, we need to create a Model object. As we want to solve our problem with constraint programming (of course, we do), we need to create a CPModel.
% ref. samples.documentation.Code4Tut1#tgettingstarted()
\lstinputlisting{java/tgetstar1.j2t}

These objects require to import the following classes:
\begin{lstlisting}
	import choco.cp.model.CPModel;
	import choco.kernel.model.Model;
\end{lstlisting}

At the begining, our model is empty, no problem has been defined explicitly. A model is composed of variables and constraints, and constraints link variables to each others.

\begin{itemize}
\item 
\textbf{Variables}\label{gettingstarted:variables}\hypertarget{gettingstarted:variables}{}
A variable is an object defined by a name, a type and a domain. We know that our unknowns are cells of the magic square. So:
\mylst{IntegerVariable cell = Choco.makeIntVar("aCell", 1, n*n);}
which means that \emph{aCell} is an integer variable, and its domain is defined from \emph{1} to \emph{n*n}.
But we need $n^2$ variables, so the easiest way to define them is:
\lstinputlisting{java/tgetstar2.j2t}

This code requires to import the following classes:
\begin{lstlisting}
	import choco.kernel.model.variables.integer.IntegerVariable;
	import choco.Choco;
\end{lstlisting}

Now that our variables are defined, we have to define the constraints between variables.
\item
\textbf{Constraints over the rows}\label{gettingstarted:constraintsovertherows}\hypertarget{gettingstarted:constraintsovertherows}{}
The sum of ach rows is equal to a constant $M$.
So we need a sum operator and and equality constraint. The both are provides by the \texttt{Choco.java} class.
\lstinputlisting{java/tgetstar3.j2t}
This part of code requires the following import:
\begin{lstlisting}
  import choco.kernel.model.constraints.Constraint;
\end{lstlisting}
After the creation of the constraints, we need to add them to the model:
\lstinputlisting{java/tgetstar4.j2t}
\item
\textbf{Constraints over the columns}\label{gettingstarted:constraintsoverthecolumns}\hypertarget{gettingstarted:constraintsoverthecolumns}{}
Now, we need to declare the equality between the sum of each column and $M$.
But, the way we have declare our variables matrix does not allow us to deal easily with it in the column case. So we create the transposed matrix (a $90^o$ rotation of the matrix) of \emph{cells}.
\begin{note}
We do not introduce new variables. We just reorder the matrix to see the \emph{column point of view}.
\end{note}
\lstinputlisting{java/tgetstar5.j2t}
Now, we can declare the constraints as before, and add them to the model:
\lstinputlisting{java/tgetstar6.j2t}
\item
\textbf{Constraints over the diagonals}\label{gettingstarted:constraintsoverthediagonals}\hypertarget{gettingstarted:constraintsoverthediagonals}{}
Now, we get the two diagonals array \emph{diags}, reordering the required \emph{cells} variables, like in the previous step.
\lstinputlisting{java/tgetstar7.j2t}

And we add the constraints to the model (in one step this time).
\lstinputlisting{java/tgetstar8.j2t}

\item
\textbf{Constraints of variables AllDifferent}\label{gettingstarted:constraintsofvariablesalldifferent}\hypertarget{gettingstarted:constraintsofvariablesalldifferent}{}
Finally, we add the AllDifferent constraints, stating that each \emph{cells} variables takes a unique value. 
One more time, we have to reorder the variables, introducing temporary array.
\lstinputlisting{java/tgetstar9.j2t}
\end{itemize}

\section{Then, the solver}\label{gettingstarted:then,thesolver}\hypertarget{gettingstarted:then,thesolver}{}
Our model is established, it does not require any other information, we can focus on the way to solve it.
The first step is to create a Solver;
\lstinputlisting{java/tgetstar10.j2t}
This part requires the following imports:
\begin{lstlisting}
	import choco.kernel.solver.Solver;
	import choco.cp.solver.CPSolver;
\end{lstlisting}

After that, the model and the solver have to be linked, thus the solver \emph{read} the model, to extract informations:
\lstinputlisting{java/tgetstar11.j2t}

Once it is done, we just need to solve it:
\lstinputlisting{java/tgetstar12.j2t}
And print the information
\lstinputlisting{java/tgetstar13.j2t}

\section{The entire code}\label{gettingstarted:thecode}\hypertarget{gettingstarted:thecode}{}
The entire code looks like:
\newpage
\lstinputlisting{java/tgetstar.j2t}

\section{Conclusion}\label{gettingstarted:conclusion}\hypertarget{gettingstarted:conclusion}{}
We saw, in a few steps, how to modelize and to solve a basic problem using constraint programming and Choco.
We think a main clue to efficiency is to strongly distinguish the modeling part from the resolution part. 
That's why, from now on, we will first introduce the modeling elements and then focus on the solving tools.   